Functional Analysis of CGM Data

### 2023-12-19 StatNet/FDA/cgm_fda.qmd
reorganize StatNet folder

---
title: "Functional Analysis of CGM Data"
author: "Tim Vigers"
format: 
  revealjs:
    scrollable: true
    smaller: true
editor: source
---

@@

```{r}
data(DTI)
DTI = subset(DTI, select = c(cca, case, pasat))
DTI = DTI[complete.cases(DTI),]
DTI$gender = factor(sample(c("male","female"), dim(DTI)[1], replace = TRUE))
DTI$status = factor(sample(c("RRMS", "SPMS", "PPMS"), dim(DTI)[1], replace = TRUE))
## fit models
dti_fosr_mod = bayes_fosr(cca ~ gender, data = DTI)
```

@@

-   Does this even seem like the correct approach?
-   Really what we want to do is classify people based on CGM tracing. Is there a good package for functional discriminant analysis?
-   FoSR in `refund` requires complete data, so how do we deal with irregular sampling across CGM tracings
    -   There is between and within-subject variability in terms length of sleep and therefore number of glucose measures
-   What does it mean that the FPCA eigenfunctions are essentially flat?
-   I'm having some trouble running the full dataset, even on a relatively beefed up desktop. Any recommendations?
-   Recommendations for easier to use packages?
    -   Data formatting is a nightmare with the various FDA packages, and unfortunately `tidyfun` is pretty limited.

>>>>

---
title: "Functional Analysis of CGM Data"
author: "Tim Vigers"
format: 
  revealjs:
    scrollable: true
    smaller: true
editor: source
bibliography: FDA.bib
csl: american-medical-association.csl
---

@@

## Clustering and classification

![](clusters.png)

- $k$-means functional clustering is the most common approach. @wangFunctionalDataAnalysis2016
  - Mean *functions* as cluster centers

## Clustering via FPCA

- Given $K$ FPCs, FPCA can be represented as:

$$
X_{iK}(t)=\mu(t)+\sum_{k=1}^K A_{ik}\phi_{k}(t)
$$
- ${A_{ik}}$ are the FPC scores and represent different cluster patterns.
- Or, if you're using a basis expansion, you can use the projections onto the basis expansion space ${B_{ik}}$
  - The advantage of FPCA is that the basis functions are determined by the data

## Classification

- Ideally, we would like to extend the clustering analysis to classification.
- Functional linear discriminant analysis would allow us to apply Bayes' rule to new CGM tracings:

$$
Pr(Z=k|X_0)=\frac{\pi_k f_k(X_0)}{\sum_{j=1}^K\pi_j f_j(X_0)}
$$

- James and Hastie (2001) @jamesFunctionalLinearDiscriminant2001 proposed a method for this, @wangFunctionalDataAnalysis2016 but even the paper is difficult to access.

@@

- Does this even seem like the correct approach? Is it worth the effort?
- Is there a good package for functional discriminant analysis? 
- FoSR in `refund` requires complete data, so how do we deal with irregular sampling across CGM tracings?
  - There is between and within-subject variability in terms length of sleep and therefore number of glucose measures
- What does it mean that the FPCA eigenfunctions are essentially flat?
- I'm having some trouble running the full dataset, even on a relatively beefed up desktop. Any recommendations?
- Recommendations for easier to use packages?
  - Data formatting is a nightmare with the various FDA packages, and unfortunately `tidyfun` is pretty limited.

## References
